Measuring causal influence with\\ back-to-back regression: the linear case

### 2019-09-25 paper/paper.tex
modifs in synthetic
synthetic
pushing things in appendix

\usepackage{amsmath, amsfonts, microtype, xcolor, tikz, graphicx, hyperref, amsthm}
\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage{iclr2020_conference,times}

\newtheorem{theorem}{Theorem}

@@

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centred and full rank noise.
     %
     If $F$ and $X$ are full-rank on $Img(E)$, then, the solution of B2B, $\hat H$, will minimize
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
\begin{proof}

 Let $\hat G$ and $\hat H$ be the solutions of the first and second regressions of B2B, we have
 \begin{align*}
    \hat G = \arg \min_G \mathbb{E}[\left \| YG - X \right \|^2] &=   \arg \min_G \mathbb{E}[\left \| X - (XE + N)FG \right\|^2]\\
%                                                        &{}= \arg \min_G \mathbb{E}[\left \| X - XEFG\right\| ^2]  + \mathbb{E}[\left \| NFG\right \| ^2]\\
                                                        &{}= \arg \min_G \left \| X - XEFG\right\| ^2  + \left \| NFG\right \| ^2
     \label{eq:doublenorm}
\end{align*}
\begin{align*}
    \hat H = \arg \min_H \mathbb{E}[\| XH - Y \hat{G} \|^2] &=\arg  \min_H \mathbb{E}[\| XH - (XE + N)F \hat G \|^2] \\
    &=\arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2] + \mathbb{E}[\| NF\hat G \| ^2]\\
%    &= \arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2]\\
    &= EF \hat G
\end{align*}

Let us prove that $EF\hat G = F\hat G$, that is, that the $d_x$ bottom rows of $F\hat G$ are zero. Let $Z=F^\dagger EF\hat G$. We have $FZ = FF^\dagger EF  \hat G= EF\hat G$ ($FF^\dagger E =E$ since $F$ is full rank on $Img(E)$). Since E is a contraction, we have $ \| NEF\hat G\|^2 \leq \| NF\hat G \|^2$. Therefore,
 $$\left \| X - XEFZ\right \| ^2  + \left \| NFZ\right \| ^2 = \| X - XEF\hat G \| ^2  + \| NEF\hat G \| ^2 \leq \| X - XEF\hat G \| ^2  + \| NF\hat G \| ^2$$

But as $\hat G =  \arg \min_G \left \| X - XEFG\right\| ^2  + \left \| NFG\right \| ^2$, the above inequality is an equality, $Z=\hat G$ and $EF\hat G = F\Hat G$. Therefore, $\left \| X - XEFG\right \| ^2  + \left \| NFG\right \| ^2 = \left \| X - XEFG\right \| ^2  + \left \| NEFG\right \| ^2 = \left \| X - XH\right \| ^2  + \left \| NH\right \| ^2$ is minimised by $\hat G$ and $\hat H$. Finally, $E\hat H = E EF\hat G = EF\hat G = \hat H$. This completes the proof.
\end{proof}

@@

>>>>

\usepackage{amsmath, amsfonts, microtype, xcolor, tikz, graphicx, hyperref, amsthm}
\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage{iclr2020_conference,times}
\DeclareMathOperator*{\argmax}{argmax} % thin space, limits underneath in displays

\newtheorem{theorem}{Theorem}
\newtheorem*{theorem*}{Theorem}

@@

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centred and full rank noise.
     %
     If $F$ and $X$ are full-rank on $Img(E)$, then, the solution of B2B, $\hat H$, will minimize
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}

\begin{proof}
  See Appendix \ref{appendix:theorem_proof}.
\end{proof}

% \begin{proof}
%
%  Let $\hat G$ and $\hat H$ be the solutions of the first and second regressions of B2B, we have
%  \begin{align*}
%     \hat G = \arg \min_G \mathbb{E}[\left \| YG - X \right \|^2] &=   \arg \min_G \mathbb{E}[\left \| X - (XE + N)FG \right\|^2]\\
% %                                                        &{}= \arg \min_G \mathbb{E}[\left \| X - XEFG\right\| ^2]  + \mathbb{E}[\left \| NFG\right \| ^2]\\
%                                                         &{}= \arg \min_G \left \| X - XEFG\right\| ^2  + \left \| NFG\right \| ^2
%      \label{eq:doublenorm}
% \end{align*}
% \begin{align*}
%     \hat H = \arg \min_H \mathbb{E}[\| XH - Y \hat{G} \|^2] &=\arg  \min_H \mathbb{E}[\| XH - (XE + N)F \hat G \|^2] \\
%     &=\arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2] + \mathbb{E}[\| NF\hat G \| ^2]\\
% %    &= \arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2]\\
%     &= EF \hat G
% \end{align*}
%
% Let us prove that $EF\hat G = F\hat G$, that is, that the $d_x$ bottom rows of $F\hat G$ are zero. Let $Z=F^\dagger EF\hat G$. We have $FZ = FF^\dagger EF  \hat G= EF\hat G$ ($FF^\dagger E =E$ since $F$ is full rank on $Img(E)$). Since E is a contraction, we have $ \| NEF\hat G\|^2 \leq \| NF\hat G \|^2$. Therefore,
%  $$\left \| X - XEFZ\right \| ^2  + \left \| NFZ\right \| ^2 = \| X - XEF\hat G \| ^2  + \| NEF\hat G \| ^2 \leq \| X - XEF\hat G \| ^2  + \| NF\hat G \| ^2$$
%
% But as $\hat G =  \arg \min_G \left \| X - XEFG\right\| ^2  + \left \| NFG\right \| ^2$, the above inequality is an equality, $Z=\hat G$ and $EF\hat G = F\Hat G$. Therefore, $\left \| X - XEFG\right \| ^2  + \left \| NFG\right \| ^2 = \left \| X - XEFG\right \| ^2  + \left \| NEFG\right \| ^2 = \left \| X - XH\right \| ^2  + \left \| NH\right \| ^2$ is minimised by $\hat G$ and $\hat H$. Finally, $E\hat H = E EF\hat G = EF\hat G = \hat H$. This completes the proof.
% \end{proof}

@@

\input{appendix}
